---
title: "plot.ly and dashboards"
author: "Zihan Xiong"
date: "2025-10-21"
output: html_document
---
# web-centric tools
plotly provides a platform for online data analytics and visualizations;
build on of HTML etc.;
"grammar of graphics" in ggplot;
convey a lot of info in a easy-to-understand format.

```{r, message=FALSE}
library(tidyverse)
library(plotly)

library(p8105.datasets)
```

Lets load the Airbnb datsets.

```{r}
data("nyc_airbnb")

nyc_airbnb=nyc_airbnb|>
  mutate(
    rating=review_scores_location/2
  )|>
  select(
    neighbourhood_group, neighbourhood, rating, price, room_type, lat, long
  )|>
  filter(
    neighbourhood_group=="Manhattan",
    room_type=="Entire home/apt",
    price <= 500,
    price >=100
  )

nyc_airbnb
```

Let's make a scatterplot but interactive at this time.

```{r}
nyc_airbnb |>
  mutate(
    label=str_c("Price:&", price, "\nRating: ", rating)
  )|>
  plot_ly(
    x = ~lat, y=~long, color=~price, 
    text=~label,
    type = "scatter", mode = "markers", alpha=.5
  )
```

lets make a boxplot
```{r}
nyc_airbnb |>
  mutate(
    neighbourhood=fct_reorder(neighbourhood,price)) |>

  plot_ly(
    x=~neighbourhood, y=~price,color=~neighbourhood,
    type="box", colors = "viridis"
  )
```

barcharts

```{r}
nyc_airbnb |> 
  count(neighbourhood) |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) |> 
  plot_ly(x = ~neighbourhood, y = ~n, color = ~neighbourhood, type = "bar", colors = "viridis")
```
## ggplotly

do not use this often!
```{r}
ggp_scatterplot=
  nyc_airbnb|>
  ggplot(aes(x=lat, y=long, color=price))+
  geom_point()

ggplotly(ggp_scatterplot)
```

# writing functions
```{r}
library(tidyverse)
library(rvest)

set.seed(1)

```
```{r}
x_vec = rnorm(25, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)
```

```{r}
z_scores = function(x) {
  
  z = (x - mean(x)) / sd(x)
  z
  
}

z_scores(x_vec)
```
```{r}
z_scores(3)
```

```{r}
z_scores("my name is jeff")
```
```{r}
z_scores(iris)
```
```{r}
z_scores(sample(c(TRUE, FALSE), 25, replace = TRUE))
```

```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = mean(x) / sd(x)
  
  z
}
```

multiple outputs
```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  list(mean = mean_x, 
       sd = sd_x)
}
```

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

multiple inputs
```{r}

```

```{r}
sim_mean_sd(mu=48,50)
```

```{r}
fellowship_ring=
  read_excel("data/LotR_Words.xlsx", range="B3:D6") |>
  mutate(movie="Fellowship of the Ring")

two_towers=
  read_excel("data/LotR_Words.xlsx", range="F3:H6") |>
  mutate(movie="Two Towers")

return_of_the_king=
  read_excel("data/LotR_Words.xlsx", range="J3:L6") |>
  mutate(movie="Return of the King")

lotr_df=
  
```

```{r}
lotr_import=function(cell_range,movie_title) {
  df=
    read_excel("data/LotR_Words.xlsx", range=cell_range) |>
    mutate(movie=movie_title)
  
  df
}

fellowship=lotr_import(cell_range="B3:D6", movie_title = "Fellowship")
two_towers=lotr_import(cell_range = "F3:H6", movie_title = "Two Towers")
return=lotr_import(cell_range="J3:L6", movie_title = "Return")

bind_rows(fellowship,two_towers,return)
```
one more example!
```{r}

```













